---
title: "DDS - Assignment 4"
author: "Linda Eliasen"
date: "February 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(XML)
library(dplyr)
library(tidyr)
library(stringi)
library(rvest)
library(ggplot2)
library(RCurl)
library(humaniformat)
library(knitr)
library(ggpubr)
theme_set(theme_pubr())
```
####*Personal SMU Repository: <https://github.com/laelias7/SMU-MSDS.git>*

```
library(XML)
library(dplyr)
library(tidyr)
library(rvest)
library(ggplot2)
library(RCurl)
library(humaniformat)
library(knitr)
library(ggpubr)
theme_set(theme_pubr())
```







##Harry Potter Cast
Scrape the full cast and crew of *Harry Potter and the Deathly Hallows: Part 2*

```{r}
hp<-read_html("http://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1")
hp_table<-html_nodes(hp,"table")
vold<-html_table(hp_table)
```

Clean up the table

```{r }
a<-data.frame(vold[3])
names(a)<-c("Blank", "Actor", "Blank2", "Character")
dhal=a[,c(2,4)]

#rename value by column, cell number
dhal[11,2]<-"Griphook/Professor Filius Flitwick" 

#remove Rest of cast listed alphabetically:
dhal <- dhal[-c(93),]
dhal<-dhal[-c(1),]

dhal$Character<-gsub("\n  \n  \n","",dhal$Character)
```

Split the Actor's name into two columns

```{r }
aname<-first_name(dhal$Actor)

bname<-middle_name(dhal$Actor)
#will need to remove NA
bname[is.na(bname)] <- ""

cname<-last_name(dhal$Actor)

fname<-paste(aname, bname)
```

Present the first 10 rows of the dataframe

```{r}
dhal2<-data.frame("First_Name"=fname, "Last_Name"=cname, "Character"=dhal$Character)
head(dhal2,10)
```

##SportsBall
Scrape the Shooting Statistics table for the San Antonio Spurs

```{r}
spurs<-read_html("http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs")
spurs_table<-html_nodes(spurs, "table")
spurs_dfs<-html_table(spurs_table, fill = TRUE)
```

Create and clean up the table.  Split the player column to separate the player's name and their position.

```{r}
Rost1 = spurs_dfs[[2]]
names(Rost1)<-c("Name")
aname<-first_name(Rost1$Name)
bname<-middle_name(Rost1$Name)
cname<-last_name(Rost1$Name)
Name<-paste(aname, bname)
Pos<-cname
Rost1<-data.frame("Name"=Name, "Position"=Pos)

Rost2 = spurs_dfs[[8]]

Rost1 <- Rost1[-c(16),]
Rost2b <- Rost2[-c(16),]

Roster<-data.frame("Name"=Rost1$Name, "Position"=Rost1$Pos, "FGM"=Rost2b$FGM, "FGA"=Rost2b$FGA,
"FGP"=Rost2b$`FG%`, "3PM"=Rost2b$'3PM', "FTM"=Rost2b$FTM, "FTA"=Rost2b$FTA, "FT%"=Rost2b$'FT%',
"2PM"=Rost2b$'2PM', "2PA"=Rost2b$'2PA', "2P%"=Rost2b$'2P%', "SC-EFF"=Rost2b$'SC-EFF', "SH-EFF"=Rost2b$'SH-EFF')

Roster
```

Confirm numerical values
```{r}
str(Roster)
```


Create a colorful bar chart that shows the Field Goals Percentage Per Game for each person

```{r}
Team<-data.frame("Name"=Rost1$Name, "Position"=Rost1$Pos, "FGP"=Rost2b$`FG%`)
Team

ggplot(Team, aes(x=Name, y=FGP, fill=Position)) +
  geom_bar(stat='identity', color='black') +
  labs(x="Player",y="Percentage per Game") +
  ggtitle("Field Goal Percentages per Game") + 
  theme_pubclean() +
  coord_flip()
```

