---
title: "DDS - Assignment 5"
author: "Linda Eliasen"
date: "February 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(knitr)
```
####*Personal SMU Repository: <https://github.com/laelias7/SMU-MSDS.git>*

```
library(dplyr)
library(tidyr)
library(knitr)
```

##Popular Baby Names
Import 2016 Nameset and display the structure and summary

```{r}
df <- read.csv("yob2016.csv")
str(df)
summary(df)
```

Name Error (ends in yyy) - finding and fixing

```{r }
grep("y{3}", df$Name, value = TRUE)
y2016<-df[- grep("Fionayyy", df$Name),]
```

Import 2015 Nameset and format

```{r }
y2015 <- read.csv("yob2015.csv")
names(y2015)<-c("Name","Sex","Count")
```

Anything interesting in the tail?
These are among the least popular boys'names in 2015.  The names are very unique, but they were still choosen!

```{r}
tail(y2015)
```

Merge two data sets with inner merge

```{r}
final<-merge(y2015,y2016, by = "Name")
final=final[,c(1,2,3,5)]
```

Add total column and sum the counts from each year

```{r}
final[,"Total"] <- NA
final$Total <- final$Count.x + final$Count.y 
names(final)<-c("Name","Sex","Count16","Count15","Total")
head(final)

```

How many people were given popular names?

```{r}
sum(final$Total)
```


What are the top 10 most popular names?

```{r}
final <- final[order(-final$Total),]
final=final[,c(1,2,5)]
head(final,10)
```

What are the top 10 most popular girls names?

```{r}
final <- final[order(final$Sex, -final$Total),] 
head(final,10)
```

Print the top ten girls' names and their counts to a csv file

```{r}
topgirls=final[,c(1,3)]
topgirls<-head(topgirls,10)
topgirls

write.csv(topgirls,'topgirls.csv')
```