---
title: "DDS Unit 03"
author: "Linda Eliasen"
date: "January 29, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(knitr)
library(ggpubr)
theme_set(theme_pubr())
```
####*Personal SMU Repository: <https://github.com/laelias7/SMU-MSDS.git>*


## GitHub Cloning

####**Clone AwesomeDataSets in GitHub**

1. https://github.com/awesomedata/apd-core.git *clone with HTTPS*

2. https://github.com/laelias7/awesome-public-datasets.git  *cloned to my repository*


####**Clone in RStudio**

- Repository URL: https://github.com/laelias7/awesome-public-datasets.git

- Project Directory Name:  AvesomeDataSets

- Create Project as subdirectory of: ~/R/Projects


*Result*
- linda@LAPTOP-9MV7J2V0  ~/Documents/R/Projects/Awesome DataSets (master)


## TITANIC

###Provide the dimensions of the data.  Count the passengers by gender and show a corresponding plot.

```{r gfile}
df<- read.csv("titanic.csv", header=TRUE)
dim(df)
```
```{r}
Gender<-xtabs(~df$Sex)
Gender
```

```{r t3}
df <- df%>%
  group_by(Sex) %>%
  summarise(counts = n())

ggplot(df, aes(x = Sex, y = counts)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) + 
  ggtitle("Pasengers by Gender") +
  theme_pubclean()
```


###Use an apply function to output the means of Age, Fare and Survival.

```{r meadt}
df<- read.csv("titanic.csv", header=TRUE)
age<-df$Age[!is.na(df$Age)]
means <- list(Age = age, Fare = df$Fare, Survival = df$Survived)
sapply(means, mean)
```



#SLEEP DATA

####Create objects for the median Age, the minimum and maximum Duration of sleep, and the mean and standard deviation of the Rosenberg Self Esteem scale (RSES).  Change the column names to MedianAge, SelfEsteem, SE_SD, and DurationRange. Run the function and show the output.


```{r sleep}
sleep01<- read.csv("Sleep_data.csv", header=TRUE)

#a=age, b=duration, c=RSES
Report <- function(a, b, c) {
  mdn <- median(a, na.rm=TRUE)
  min <- min(b, na.rm=TRUE)
  max <- max(b, na.rm=TRUE)
  mea <- mean(c, na.rm=TRUE)/5
  std <- sd(c, na.rm=TRUE)/5
  range <- max-min
  Function <- c("MedianAge", "SelfEsteem", "SE_SD", "DurationRange")
  Report <- round(c(mdn, mea, std, range),2)
  result <- data.frame(Function,Report)
  result
}

Report(a=sleep01$Age, b=sleep01$Duration, c=sleep01$RSES)
```


#FIVETHIRTYEIGHT

```{r cldta}
data(package = "fivethirtyeight")
df<- fivethirtyeight::college_recent_grads
```

#####https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/


####Get the dimensions,column names and column count of the dataset
```{r dfive}
df<- fivethirtyeight::college_recent_grads
dim(df)
names(df)
ncol(df)
```

#### Generate a count for each unique major category and assign it to major_count.  Plot the resulting data.

```{r m}
major_count<-data.frame(xtabs(~df$major_category))
major_count
```

```{r}
df <- df %>%
  group_by(major_category) %>%
  summarise(counts = n())

ggplot(df, aes(x = major_category, y = counts)) +
  geom_bar(fill = "#0073C2FF", stat = "identity") +
  geom_text(aes(label = counts), vjust = -0.3) +
  ggtitle("Number of Degrees by Category") +
  theme_pubclean() +
  coord_flip()
```